# Benford Analysis

## Create Benford object

We use the [benford.analysis package](https://www.rdocumentation.org/packages/benford.analysis/versions/0.1.5), running its `benford` function against the BTC_master data, producing a Benford object, which we'll call `bfd.mast`. From `bfd.mast`, we will extract information to perform a Benford analysis on the `Vol_BTC` column only. The thinking is that volumes are an indicator of market liquidity and, therefore, a measure of the reality of Bitcoin and, therefore, the viability of MSTR as a going concern.

## Benford plots

Applying the `plot` function to `bfd.mast` generates a prearranged set of plots, including several that are not technically Benford. These can offer interesting insights on the data and may point an auditor toward groups of observations that may merit further scrutiny. 

```{r}
plot(bfd.mast)
```

The Digits Distribution plot shows a few spikes above Benford's Law expectations in the 10s, 20s, and 50s. These deserve a closer look, despite the MAD "Close Conformity" rating. Let's zoom in with homemade visualization.

```{r}
# Extract data from bfd.mast
KurtBfd <- data.frame(as.character(bfd.mast$bfd$digits), 
                    bfd.mast$bfd$data.dist, 
                    bfd.mast$bfd$benford.dist) 
# Rename columns
KurtBfd <- KurtBfd |> 
  rename(Digits = as.character.bfd.mast.bfd.digits., 
         Actual = bfd.mast.bfd.data.dist,        
         BfdDist = bfd.mast.bfd.benford.dist)

# Plot distribution
ggplot(KurtBfd) +   
  geom_line(aes(Digits, BfdDist), group = 1, color = "blue", linewidth = .5, linetype = 2) +     geom_bar(aes(Digits, Actual), stat = "identity", fill = "darkgreen", alpha = 0.5)+   
  theme_minimal()+   
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust = 1, hjust = 0)) +
  scale_x_discrete(breaks = seq(10, 90, by = 10))+
  labs(title = "Distribution of First-Two Digits Pairs", subtitle = "Bitcoin vs. Benford: 2011-2021 ", y = "Relative frequency", x = "First two digits")
```

This more granular view brings into focus spikes on the tens, which suggests either a market-specific artefact or manipulation. Either way, the drill down on the volumes.

The Summation Distribution and Difference plots – which are not technically Benford's Law tests – appear to support detailed tests of individual observations. This is because of the obvious discrepancies between the horizontal red line and the height of the digit-pair bars. However, interpretation of these plots is complicated by lack of empirical support for an expected distribution of sums across the pairs. No one yet knows how high each of the bars should be. 

Nigrini suggests further testing the bars that are highest in their neighborhoods.[^07-benford-1] Given that the summation test naturally segregates observations into bins containing differing volumes of units (e.g., currency), the summation test seems a good departure point for probability-proportional-to-size (PPS) sampling.

[^07-benford-1]: Mark Nigrini, *Forensic Analytics: Methods and Techniques for Forensic Accounting Investigations*, 2e (2020), p. 92.

The remaining plots are of limited usefulness. The Second Order plot is questionable because "in general, there is no relationship between accounting data and the differences between the ordered/ranked values in a data set."[^07-benford-2] The Chi-Squared Difference plot should be discounted because the sample size is so large, as explained in Chapter \@ref(context). Having reviewed the plots, it's time to drill down with some queries against `bfd.mast`.

[^07-benford-2]: Nigrini (2020) at 92.


```{r}

bfd.mast <- benford(BTC_master$Vol_BTC, 2, sign="positive")  
bfd.mast

```

## Interpret `benford` output

This output presents a lot to absorb, but we'll keep it simple. One statistic of note is the MAD (mean absolute deviation). Based on the significance thresholds proposed by [@benford2012], which are programmed into the benford.analysis package[@benford.analysis], the MAD of 0.001091855 suggests little overall deviation from Benford's Law, hence the "Close Conformity" rating. However, the information value of MAD and the other statistics presented is open to question for reasons explained in Chapter \@ref(context). he plots, which reveal some interesting patterns.

## Drill down

The Benford object, `bfd.mast`, contains a wealth of additional information. Using `benford.analysis` functions, we will access and comment on some of this information below.

### Top ten suspects

Here, we create a table of the top ten "suspect" *digits pairs* sorted by decreasing absolute deviation from the frequencies predicted by Benford's Law.

```{r}

head(suspectsTable(bfd.mast),10)

```

### Get suspects

Given that pairs 10 and 20 are the top two suspects, `getSuspects` with `how.many=2` retrieves all *observations* that begin with 10 or 20, sorted by ascending Timestamp. There are 325148 such observations. The `head` function limits the output shown to only the first six.

```{r Suspect obs}

head(getSuspects(bfd.mast, BTC_master, how.many=2))

```

### Duplicate values

`duplicatesTable` produces a table of duplicated Vol_BTC *values* sorted by number of duplicates (descending).

```{r duplicatesTable}

duplicatesTable(bfd.mast)

```

Here, we call *observations* with the two most frequently duplicated Vol_BTC values.

```{r getDuplicates}

duplicates <- getDuplicates(bfd.mast, BTC_master, how.many = 2) 
head(duplicates)

```

Get observations starting with pairs 50 or 99.

```{r 50 or 99}

digits_50_and_99 <- getDigits(bfd.mast, BTC_master, digits=c(50, 99))
head(digits_50_and_99)

```

## EMAD and Excess MAD

A significant drawback of Benford's Law is the lack of objectively proven measures of conformity. What is a material departure from Benford's expectations? While Nigrini has proposed some broad thresholds, they remain largely without statistical support.

One of the greatest challenges is that reliable Benford's Law readings are obtainable only with Ns (sample sizes) in the thousands. With Ns so large, traditional statistics applied to MAD values – e.g., Z scores, Chi-square, and Kolmogorov-Smirnov – throw false positives because of excess power.[^07-benford-3] In fact, MAD scores and N are so strongly correlated that the linear relationship between them can be used to estimate "expected MAD" (or EMAD), which is the portion of actual MAD attributable to chance. EMAD is a function of N as stated in Equation 2.

[^07-benford-3]: See Bradley Barney & Kurt Schulzke (2016), Moderating ‘‘Cry Wolf’’ Events with Excess MAD in Benford’s Law Research and Practice, A66, A70. *J. Forensic Accounting Research*; Vol. 1 No. 1, DOI: 10.2308/jfar-51622; Mark Nigrini (2017), Audit Sampling Using Benford’s Law: A Review of the Literature with Some New Perspectives, *J. Emerging Technologies in Accounting*, Vol. 14 No. 2, pp. 30-31, 40. DOI: 10.2308/jeta-51783.

`Equation 2` EMAD = 1/sqrt(158.8\*n)

EMAD gives rise to the concept of "excess" MAD, or MAD in excess of MAD attributable to chance. Any actual MAD not in excess of EMAD is "expected" and therefore, per se, justifies no suspicion of number manipulation. Excess MAD is as given in Equation 3.

`Equation 3` ExcessMAD = Actual MAD - EMAD

These numbers are calculated and reported in the following chunk.

```{r EMAD etc}

EMAD <- 1/sqrt(158.8*bfd.mast$info$n) 
ExcessMAD <- bfd.mast$MAD - EMAD 
bfd.mast$MAD 
EMAD 
ExcessMAD
ExcessMAD/bfd.mast$MAD
```

So ExcessMAD is a very small positive number positive number, approximately equal to MAD. Unfortunately, as with MAD itself, how much Excess MAD is "too much" remains speculative. However, we can say that zero Excess MAD is definitively too little to raise an alarm over manipulated data. 

## Benford's Law Conclusion

This Benford analysis focused exclusively on 4857377 per-minute purported Bitcoin transaction volume values drawn from a Kaggle data set said to cover the 2011 Q4 - 2021 Q1 period. The tables and visualizations produced
