# My Law analysis

The My Law concept was first articulated in print by Mark Nigrini in 2012. The idea is that individual organizations may develop, in their data sets, first-digits patterns that do not conform to Benford's Law but are internally consistent over time. Thus, these proprietary digital patterns might be used as a "My Law" in place of Benford's Law. Significant departures from My Law might call for further investigation.[^07-mylaw-1]

[^07-mylaw-1]: Mark Nigrini, Benford's Law: Applications for Forensic Accounting, Auditing, and Fraud Detection (2012), pp. 295-300.

In the current exercise, we'll use Kaggle's BTC volumes numbers for the period 2018-2020 to establish the My Law distribution benchmark. We'll then compare 2021 volume numbers to the benchmark, calculate a MAD statistic and discuss the implications.

## Benford x 2

Similar to the Benford analysis above, our first step is to use `benford` to create the necessary Benford objects for 2018-2020 and 2021. We'll then use information contained in these objects to create a separate MyLaw object.

```{r}

bfd.18_20 <- benford(BTC_18_20$Vol_BTC) 
bfd.21 <- benford(BTC_21$Vol_BTC)

```

## Create MyLaw object

The MyLaw object will contain four columns. Run this chunk and then use `head` to view what's inside.

```{r}

MyLaw <- data.frame(as.character(bfd.18_20$bfd$digits), 
                    bfd.18_20$bfd$data.dist, 
                    bfd.21$bfd$data.dist, 
                    bfd.21$bfd$benford.dist) 
head(MyLaw)
```

We'll now rename the MyLaw columns for ease in writing (and later reading) the script.

```{r}
MyLaw <- MyLaw |>    
  
  rename(Digits = as.character.bfd.18_20.bfd.digits., 
         MyLaw = bfd.18_20.bfd.data.dist,        
         Vol_21_actual = bfd.21.bfd.data.dist, 
         BfdDist = bfd.21.bfd.benford.dist)
```

Here, we create a Diffs column in MyLaw.

```{r MyLaw diffs}
MyLaw$Diffs <- MyLaw$Vol_21_actual - MyLaw$MyLaw 
```

### My Law Plots

The first plot shows -- for minute-by-minute transaction volumes -- the actual distribution of first-two digits in the Kaggle data (green histogram); the Benford's Law predicted distribution (red dashed line); and the My Law distribution (blue dashed line) derived from Kaggle data for 2018-2020.[^07-mylaw-2]

[^07-mylaw-2]: Michael Clark's Benford's ADA analysis submission in ACCT 7530 contributed some snippets and inspired others used in the My Law analysis.

```{r MyLaw plots, fig.align='center', message=FALSE, warning=FALSE}

MyLawPlot <- ggplot(MyLaw) +   
  geom_line(aes(Digits, MyLaw), group = 1, color = "blue", linewidth = 1, linetype = 2) +   
  geom_bar(aes(Digits, Vol_21_actual), stat = "identity", fill = "darkgreen", alpha = 0.5) +   
  geom_line(aes(Digits, BfdDist), group = 1, color = "red", linetype = 2, linewidth = 1)  

MyLawPlot +   
  theme_minimal()+   
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 1, hjust = 1.5)) +   
  labs(y = "Relative frequency", x = "First two digits")
```

## Plot diffs by pairs

Here, we plot as a bar chart the *differences* between the actual and My Law frequencies of each pair of first digits. The "fill" color is implemented with `scale_fill_gradient`, with blue as the low end and red as the high. Experiment with alternative high and low colors to see what happens.     

```{r Plot MyLaw diffs, fig.align='center', message=FALSE, warning=FALSE}

ggplot(MyLaw) + 
  geom_bar(aes(Digits, Vol_21_actual - MyLaw, fill = Vol_21_actual -  MyLaw), stat = "identity", show.legend = FALSE) +   
  scale_fill_gradient(low = "blue", high = "red") +   
  theme_minimal() +   
  theme(axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1)) +   
  labs(y = "Difference in relative frequency", x = "First two digits")
```

Perhaps the biggest taken from the gradients is how little the actual frequencies deviate from My Law. This at least suggests that these data do have their own "native" first-digits pattern.

### Calculate MyLaw MAD

```{r MyLaw MAD}

MyLawMAD <- mean(abs(MyLaw$Vol_21_actual - MyLaw$MyLaw)) 
MyLawMAD 

```

## My Law Conclusion

[Evaluate the results of the ADA]
